Probabilities and propensities: conversion experiences in online grocery shopping

### 2017-07-19 log.rmd
rewriting conclusion

>>>>

## Wed 19 Jul 2017 06:46:17 BST
- try writing the conclusion to give shape to various shifts in emphasis;  
- made more notes in Dewey on reconstruction; see dewey_reconstruction_1957.md
- TODO: put more Dewey in intro and conclusion

- order of points in the conclusion:
    - the specific situation is grocery list  and online shopping recommendations, not shopping in general, but the conversion or transformation of a  recommender system. 
    - a small theory of shopping and shopping lists: shopping as a constant negotiation between normative and actual orders; list as an intersectional ordering device -- it creates a set of associations; it maps them to an order; but that order can be changed and can be re-associated
    - these systems largely understood in terms of personalization; undoubtedly occurring. But  something else underpins  personalization and allows it to occur
    - probabilization allows personalization and exceeds it: associations seen in the recommender system include person-thing, thing-thing, person-in-time; these associations overflow closure; the argument here has focused on describing how this affects attempts to construct predictive models
    - suggests an alternative theory of social order in which association is not any single organization but propensities, or tendencies towards realization; 
    - conversion events are what is going on in big data more generally: how do we think about them without being caught in inherited patterns of thought. 
    - reconstruction and what it does: focuses on specific situations in order to get at the trouble, the problem, the conflict that attracts our interest to them; specifically addresses a broader inherited problem in relation between science and society

- ok, that might work, but still a lot happening; there is no single clear point here at the moment; will have to simplify this more to make it work as article; the point that matters most to me is the personalization one; but I'm not sure exactly why ... I know philosophically -- I'm interested in probability -- but not so sure why in terms of the argument of the paper;   using probability and 
- now trying to draft the conclusion: wrote 500 words;

### 2017-07-19 reconstruction.rmd
rewriting conclusion

>>>>

The 'minute and extensive scrutinizing' [@Dewey_1957] of a recommender system undergoing a conversion from an older data-mining approach to a current 'personal relevance model' approach addresses a specific situation: online grocery shopping (not shopping in general) and the ordering of lists.

The theory of shopping lists constructed at the beginning of the paper was a scaffolding component. It draws on anthropology, sociology and mathematics to suggest that we see shopping lists as the site of a complex negotiation between different social orders. The list itself as an inscriptive device functions as an ordering device, a device that accommodates an immense variety of specific habits, dispositions, values and associations. The mapping component of the list, the order that arranges the set of items according to different interests and aims, is the operational focus of recommender systems.

The personalization of recommendations has been a defining characteristic of many big data conversion events. Rightly, the effects of all-consuming personalization have been the target of critics and critical research. The archaeology of the Tesco recommender system, including its early experiments with demographic data-mining, its adoption of a rule-based 'market basket' analysis system, and its current implementation of a 'personal relevance model' certainly follows that trajectory.   But the conversion event is not reducible to personalization.

Probabilization precedes personalization. When we observe what is practically done to construct a personal relevance model and apply it to a shopping list, we see, conditional probabilities, predictive models, the importance of data accumulated over time,  platform experiences, the rhythms of infrastructural maintenance, and open-ended construction of new features in data, all of which negotiate shifting and complicated associations between people and things, people and people, and things with things. If social order is made of associations, is to be social is to associate, then the forms of association taking shape in this and other similar conversion events might be understood as understand as operationalizing probability in matrices of propensity. As Popper argues, probabilities are being realized in the same way that fields of force were realized in modern physics. Probabilization does not take place in physics laboratories but on platforms, combining databases, compute capacity, interfaces and global supply chain logistics.

The fact that probabilisation largely takes the form of personalization tends to obscure the re-framing of social order taking place. This ordering  depends on  the mappings that inscriptive devices such as shopping lists  display. Only because shopping is already somewhat ordered by lists can a partial re-mapping, provisionally conducive to new events, take place. The conversion entailed in this re-mapping is not just about perosnalization. It is more ontological than that since it includes relations running multi-laterally between people and things in time. It is less ontological, however, in that it is not a sudden transformation, but an ongoing process of adjusting, scaffolding, intervening and configuring orders.

Reconstructions of specific conversion events associated with big data can work in different ways. The combination of social theory, archaeology, philosophy and model/data prototypes used in this paper sketches one possible way of addressing the trouble, the disturbance or instability that seems to affect many discussions of big data.

------------
